SUPPLEMENTARY MATERIALS for\\ ~\\ \textbf{Generalized additive models to analyze biomedical non-linear longitudinal data in R:}\\ Beyond repeated measures ANOVA and Linear Mixed Models\\ ~\\ APPENDIX B: CODE

### 2021-12-29 Manuscripts/Manuscript_by_chapters-SIM_Revisions_final/appendices/Appendix_B.tex
revision to appendix

%\lstset{
%  breaklines=true,
%  stringstyle=\ttfamily,
%  backgroundcolor=\color{gray}
%}
\usepackage{placeins}
\usepackage{subfig}
\usepackage{breqn}
\usepackage[font={small}]{caption}
\usepackage{float}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{SUPPLEMENTARY MATERIALS for\\
\strut \\
\textbf{Generalized additive models to analyze biomedical non-linear longitudinal data in R:}\\
Beyond repeated measures ANOVA and Linear Mixed Models\\
\strut \\
APPENDIX B: CODE}
\author{}
\date{\vspace{-2.5em}}

@@

#alpha for ribbon in the smooth plots
al<-0.8

thm1<-scale_fill_scico_d(palette="tokyo",begin=0.3, end=0.8, direction = -1, aesthetics = c("colour","fill"))
\end{lstlisting}

@@

The first function is \passthrough{\lstinline!example.R!}, which allows to simulate linear and quadratic data in the same manner as in Section 3.5 in the main manuscript. Both rm-ANOVA and LMEM with interaction are fitted to the data. The error for each simulated trend can be correlated or uncorrelated as well.

@@

  #return everything in a list
  return(list(
    dat = dat,
    pred_dat = pred_dat,
    fit_anova=fit_anova,
    fit_lme = fit_lme
  ))
}
\end{lstlisting}

@@

Function \passthrough{\lstinline!plot\_example.R!} uses the output of \passthrough{\lstinline!example.R!} to show the fit of a rm-ANOVA and a LMEM. This function can be used to show an expanded version of Figure 1 in the main manuscript, presenting simulated data with correlated and uncorrelated errors and how the individual trends vary in each case. The corresponding rm-ANOVA and LMEM fits are also presented, and we show the complete output in the next subsection.

\begin{lstlisting}[language=R]
## This function plots the rm-ANOVA and LMEM for the data simulated in example.R
plot_example <- function(sim_dat) {
    txt<-20
    p1 <- sim_dat$dat %>%
        ggplot(aes(x = time,
                   y = y,
                   group = treatment,
                   color = treatment)
        ) +
        geom_point(show.legend=FALSE) +
        labs(y='response')+
        geom_line(aes(x = time,
                      y = mu,
                      color = treatment),
                  show.legend=FALSE) +
        theme_classic() +
        theme(plot.title = element_text(size = txt,
                                        face = "bold"),
              text=element_text(size=txt))+
        thm1

    #plot the simulated data with trajectories per each subject
    p2 <- sim_dat$dat %>%
        ggplot(aes(x = time,
                   y = y,
                   group = subject,
                   color = treatment)
        ) +
        geom_line(aes(size = "Subjects"),
                  show.legend = FALSE) +
        # facet_wrap(~ treatment) +
        geom_line(aes(x = time,
                      y = mu,
                      color = treatment,
                      size = "Simulated Truth"),
                  lty = 1,show.legend = FALSE) +
        labs(y='response')+
        scale_size_manual(name = "Type", values=c("Subjects" = 0.5, "Simulated Truth" = 3)) +
        theme_classic()+
        theme(plot.title = element_text(size = txt,
                                        face = "bold"),
              text=element_text(size=txt))+
        thm1

    #plot the errors
    p3 <- sim_dat$dat %>%
        ggplot(aes(x = time,
                   y = errors,
                   group = subject,
                   color = treatment)) +
        geom_line(show.legend=FALSE) +
        labs(y='errors')+
        theme_classic()+
        theme(plot.title = element_text(size = txt,
                                        face = "bold"),
              text=element_text(size=txt))+
        thm1

    #plot the model predictions for rm-ANOVA
    p4 <- ggplot(sim_dat$dat,
                 aes(x = time,
                     y = y,
                     color = treatment)) +
        geom_point(show.legend=FALSE)+
        labs(y='response')+
        geom_line(aes(y = predict(sim_dat$fit_anova),
                      group = subject, size = "Subjects"),show.legend = FALSE) +
        geom_line(data = sim_dat$pred_dat,
                  aes(y = predict(sim_dat$fit_anova,
                                  level = 0,
                                  newdata = sim_dat$pred_dat),
                      size = "Population"),
                  show.legend=FALSE) +
        guides(color = guide_legend(override.aes = list(size = 2)))+
        scale_size_manual(name = "Predictions",
                          values=c("Subjects" = 0.5, "Population" = 3)) +
        theme_classic() +
        theme(plot.title = element_text(size = txt,
                                        face = "bold"),
              text=element_text(size=txt))+
        thm1

    #plot the LMEM predictions
    p5 <- ggplot(sim_dat$dat,
                 aes(x = time,
                     y = y,
                     color = treatment)) +
        geom_point()+
        labs(y='response')+
        geom_line(aes(y = predict(sim_dat$fit_lme),
                      group = subject, size = "Subjects")) +
        geom_line(data = sim_dat$pred_dat,
                  aes(y = predict(sim_dat$fit_lme,
                                  level = 0,
                                  newdata = sim_dat$pred_dat),
                      size = "Population")) +
        guides(color = guide_legend(override.aes = list(size = 2)))+
        scale_size_manual(name = "Predictions",
                          values=c("Subjects" = 0.5, "Population" = 3)) +
        theme_classic() +
        theme(plot.title = element_text(size = txt,
                                        face = "bold"),
              text=element_text(size=txt))+
        thm1

    if(option=='simple'){
        return((p1+p4+p5)+plot_layout(nrow=1)+plot_annotation(tag_levels = 'A'))
    }
    else {
    return((p1+p3+p2+p4+p5)+plot_layout(nrow=1)+plot_annotation(tag_levels = 'A'))
    }

@@

The chunk below calls both \passthrough{\lstinline!example.R!} and \passthrough{\lstinline!plot\_example\_Appendix.R!} to simulate data and create the composite plots.

\begin{lstlisting}[language=R]
source(here::here("Manuscripts/Manuscript_by_chapters-SIM_Revisions_final/scripts","example.R"))
source(here::here("Manuscripts/Manuscript_by_chapters-SIM_Revisions_final/scripts","plot_example.R"))

A1<-plot_example(example(fun_type = "linear", error_type = "correlated"),option='composite')

B1<-plot_example(example(fun_type = "linear", error_type = "independent"),option='composite')

C1<-plot_example(example(fun_type = "quadratic", error_type = "correlated"),option='composite')

D1<-plot_example(example(fun_type = "quadratic", error_type = "independent"),option='composite') 
\end{lstlisting}

@@

\begin{lstlisting}[language=R]
#basis functions: this script creates Fig 2. by calculating a GAM for the Group 1 data of Figure 1,
# extracting the basis functions
#and creates objects p11,p12,p13,p14 for plotting, which are combined in b_plot to create the
#final composite figure
n_time = 6
x <- seq(1,6, length.out = n_time)
mu <- matrix(0, length(x), 2)
mu[, 1] <-  -(0.25 * x^2) +1.5*x-1.25 #mean response
mu[, 2] <- (0.25 * x^2) -1.5*x+1.25 #mean response
y <- array(0, dim = c(length(x), 2, 10))
errors <- array(0, dim = c(length(x), 2, 10))
for (i in 1:2) {     # number of treatments
    for (j in 1:10) {  # number of subjects
        # compound symmetry errors
        errors[, i, j] <- rmvn(1, rep(0, length(x)), 0.1 * diag(6) + 0.25 * matrix(1, 6, 6))
        y[, i, j] <- mu[, i] + errors[, i, j]
    }
}

@@

smooth<-basis_plot%>%
    group_by(time)%>%
    summarize(smooth=sum(mod_val))

#original basis
sz<-1
p11<-ggplot(basis_plot,
            aes(x=time,
                y=value_orig,
                colour=as.factor(Basis)
            )
)+
    geom_line(size=sz,
              show.legend=FALSE)+
    geom_point(size=sz+1,
               show.legend = FALSE)+
    labs(y='Basis functions')+
    facet_wrap(~Basis,
               labeller = as_labeller(basis_names)
    )+
    theme_classic()+
    thm1

#penalized basis
p12<-ggplot(basis_plot,
            aes(x=time,
                y=mod_val,
                colour=as.factor(Basis)
            )
)+
    geom_line(show.legend = FALSE,
              size=sz)+
    geom_point(show.legend = FALSE,
               size=sz+1)+
    labs(y='Penalized \n basis functions')+
    scale_y_continuous(breaks=seq(-1,1,1))+
    facet_wrap(~Basis,
               labeller=as_labeller(basis_names)
    )+
    theme_classic()+
    thm1

#heatmap of the  coefficients
x_labels<-c("Intercept","1","2","3","4")
p13<-ggplot(basis_plot,
            aes(x=Basis,
                y=Basis))+
    geom_tile(aes(fill = cof),
              colour = "black") +
    scale_fill_gradient(low = "white",
                        high = "#B50A2AFF")+
    labs(x='Basis',
         y='Basis')+
    scale_x_discrete(labels=x_labels)+
    geom_text(aes(label=round(cof,2)),
              size=7,
              show.legend = FALSE)+
    theme_classic()+
    theme(legend.title = element_blank())

#plotting simulated datapoints and smooth term
p14<-ggplot(data=dat,
            aes(x=time,y=y))+
    geom_point(size=sz+1,alpha=0.5)+
    thm1+
    labs(y='Simulated \n response')+
    geom_line(data=smooth,
              aes(x=time,
                  y=smooth),
              color="#6C581DFF",
              size=sz+1)+
    theme_classic()

#Combining all
b_plot<-p11+p13+p12+p14+plot_annotation(tag_levels='A')&
    theme(
        text=element_text(size=18)
    )
\end{lstlisting}

@@

The next chunk presents the code of the function \passthrough{\lstinline!simulate\_data.R!}, which we use in the main manuscript to simulate data that follows the reported trends in \(\mbox{StO}_2\) in the literature. Although the default standard deviation (SD) to generate the normally-distributed data in the function is 10\%, we use a standard deviation of 10\% in the main manuscript. We use the default number of 10 observations at each time point in the simulated data. Note that the input for the function is \passthrough{\lstinline!dat!}, needs to be a dataframe that contains the original values for the trends in each group over time from the literature. In our manuscript, we use the dataframe

\begin{lstlisting}[language=R]
#data from Vishwanath et. al.
dat<-tibble(StO2=c(4,27,3,2,0.5,7,4,50,45,56),
            Day=rep(c(0,2,5,7,10),times=2),
            Group=as.factor(rep(c("Control","Treatment"),each=5))
            )
\end{lstlisting}

@@

\begin{lstlisting}[language=R]
#the same model used for the full dataset
mod_m1 <- gam(StO2_sim ~ Group+s(Day,by=Group,k=5), data  = dat_missing)
#appraise the model
appraise(mod_m1)

@@

f6<-ggplot(ci, aes(x = Day, y = est, group = smooth)) +
geom_line(lwd = 1) +
geom_ribbon(data = ci, mapping = aes(ymin = lower, ymax = upper, x = Day, group = smooth,fill = Group),
inherit.aes = FALSE, alpha = 0.7,
show.legend=FALSE) +
geom_ribbon(data = si,
mapping = aes(ymin = lower, ymax = upper, x = Day, group = smooth,fill =Group),
inherit.aes = FALSE, alpha = 0.4,
show.legend=TRUE)+
    geom_line(data=si,aes(Day,upper,color=Group), size=0.8, alpha=0.7)+
    geom_line(data=si,aes(Day,lower,color=Group), size=0.8, alpha=0.7)+
    geom_point(data=dat_missing, aes(x = Day, 
                        y = StO2_sim, 
                        color = Group), 
                            size=1.5,
                        alpha=0.6, 
               inherit.aes = FALSE)+
    labs(y=expression(atop(StO[2],'(incomplete observations)')))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
      theme_classic()+
  theme(
    axis.text=element_text(size=22))+
        thm1
\end{lstlisting}

@@

Then, \passthrough{\lstinline!difference\_smooths.R!} takes the output from the previous function and uses it to compute the simultaneous CI.
We present the code below.

@@

\begin{lstlisting}[language=R]
#Function that uses the source code of `gam.check` to obtain the estimates without the plots. The source can be checked by typing `gam.check` in the console.
gam_diagnostics<-function (b, old.style = FALSE, type = c("deviance", "pearson",
                                                          "response"), k.sample = 5000, k.rep = 200, rep = 0,
                           level = 0.9, rl.col = 2, rep.col = "gray80", ...)
{
    type <- match.arg(type)
    resid <- residuals(b, type = type)
    linpred <- if (is.matrix(b$linear.predictors) && !is.matrix(resid))
        napredict(b$na.action, b$linear.predictors[, 1])
    else napredict(b$na.action, b$linear.predictors)

    fv <- if (inherits(b$family, "extended.family"))
        predict(b, type = "response")
    else fitted(b)
    if (is.matrix(fv) && !is.matrix(b$y))
        fv <- fv[, 1]
    gamm <- !(b$method %in% c("GCV", "GACV", "UBRE",
                              "REML", "ML", "P-ML", "P-REML",
                              "fREML"))
    if (gamm) {
        cat("\n'gamm' based fit - care required with interpretation.")
        cat("\nChecks based on working residuals may be misleading.")
    }
    else {
        cat("\nMethod:", b$method, "  Optimizer:",
            b$optimizer)
        if (!is.null(b$outer.info)) {
            if (b$optimizer[2] %in% c("newton", "bfgs")) {
                boi <- b$outer.info
                cat("\n", boi$conv, " after ", boi$iter,
                    " iteration", sep = "")
                if (boi$iter == 1)
                    cat(".")
                else cat("s.")
                cat("\nGradient range [", min(boi$grad),
                    ",", max(boi$grad), "]", sep = "")
                cat("\n(score ", b$gcv.ubre, " & scale ",
                    b$sig2, ").", sep = "")
                ev <- eigen(boi$hess)$values
                if (min(ev) > 0)
                    cat("\nHessian positive definite, ")
                else cat("\n")
                cat("eigenvalue range [", min(ev), ",",
                    max(ev), "].\n", sep = "")
            }
            else {
                cat("\n")
                print(b$outer.info)
            }
        }
        else {
            if (length(b$sp) == 0)
                cat("\nModel required no smoothing parameter selection")
            else {
                cat("\nSmoothing parameter selection converged after",
                    b$mgcv.conv$iter, "iteration")
                if (b$mgcv.conv$iter > 1)
                    cat("s")
                if (!b$mgcv.conv$fully.converged)
                    cat(" by steepest\ndescent step failure.\n")
                else cat(".\n")
                cat("The RMS", b$method, "score gradient at convergence was",
                    b$mgcv.conv$rms.grad, ".\n")
                if (b$mgcv.conv$hess.pos.def)
                    cat("The Hessian was positive definite.\n")
                else cat("The Hessian was not positive definite.\n")
            }
        }
        if (!is.null(b$rank)) {
            cat("Model rank = ", b$rank, "/", length(b$coefficients),
                "\n")
        }
    }
    cat("\n")
    kchck <- k.check(b, subsample = k.sample, n.rep = k.rep)
    if (!is.null(kchck)) {
        cat("Basis dimension (k) checking results. Low p-value (k-index<1) may\n")
        cat("indicate that k is too low, especially if edf is close to k'.\n\n")
        printCoefmat(kchck, digits = 3)
    }
}
\end{lstlisting}

>>>>

%\lstset{
%  breaklines=true,
%  stringstyle=\ttfamily,
%  backgroundcolor=\color{gray}
%}
\usepackage{placeins}
\usepackage{subfig}
\usepackage{breqn}
\usepackage[font={small}]{caption}
\usepackage{float}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{SUPPLEMENTARY MATERIALS for\\
~\\
\textbf{Generalized additive models to analyze biomedical non-linear longitudinal data in R:}\\
Beyond repeated measures ANOVA and Linear Mixed Models\\
~\\
APPENDIX B: CODE}
\author{}
\date{\vspace{-2.5em}}

@@

#alpha for ribbon in the smooth plots
al<-     0.8

thm1 <- scale_fill_scico_d(palette = "tokyo", begin = 0.3, end = 0.8,
                           direction = -1, aesthetics = c("colour", "fill"))
\end{lstlisting}

@@

The first function is \passthrough{\lstinline!example()!} which is in the file named \passthrough{\lstinline!example.R!} in the \passthrough{\lstinline!scripts/!} folder, which allows to simulate linear and quadratic data in the same manner as in Section 3.5 in the main manuscript. Both rm-ANOVA and LMEM with interaction are fitted to the data. The error for each simulated trend can be correlated or uncorrelated as well.

@@

  #return everything in a list
  return(list(
    dat       = dat,
    pred_dat  = pred_dat,
    fit_anova = fit_anova,
    fit_lme   = fit_lme
  ))
}
\end{lstlisting}

@@

Function \passthrough{\lstinline!plot\_example()!} from the file \passthrough{\lstinline!plot\_example.R!} in the folder \passthrough{\lstinline!scripts/!} uses the output of \passthrough{\lstinline!example.R!} to show the fit of a rm-ANOVA and a LMEM. This function can be used to show an expanded version of Figure 1 in the main manuscript, presenting simulated data with correlated and uncorrelated errors and how the individual trends vary in each case. The corresponding rm-ANOVA and LMEM fits are also presented, and we show the complete output in the next subsection.

\begin{lstlisting}[language=R]
## This function plots the rm-ANOVA and LMEM for the data simulated in example.R
plot_example <- function(sim_dat) {
  txt <- 20
  p1 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = y, group = treatment, color = treatment)) +
    geom_point(show.legend = FALSE) +
    labs(y = 'response')+
    geom_line(aes(x = time, y = mu, color = treatment), 
              show.legend = FALSE) +
    theme_classic() +
    theme(plot.title = element_text(size = txt, face = "bold"),
          text=element_text(size = txt)) +
    thm1

  #plot the simulated data with trajectories per each subject
  p2 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = y, group = subject, color = treatment)) +
    geom_line(aes(size = "Subjects"), show.legend = FALSE) +
    # facet_wrap(~ treatment) +
    geom_line(aes(x = time, y = mu, color = treatment, 
                  size = "Simulated Truth"),
              lty = 1, show.legend = FALSE) +
    labs(y = 'response') +
    scale_size_manual(name = "Type", 
                      values = c("Subjects" = 0.5, "Simulated Truth" = 3)) +
    theme_classic() +
    theme(plot.title = element_text(size = txt, face = "bold"),
          text = element_text(size = txt)) +
    thm1

  #plot the errors
  p3 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = errors, group = subject, color = treatment)) +
    geom_line(show.legend = FALSE) +
    labs(y = 'errors') +
    theme_classic() +
    theme(plot.title = element_text(size = txt, face = "bold"),
          text = element_text(size = txt))+
    thm1

  #plot the model predictions for rm-ANOVA
  p4 <- ggplot(sim_dat$dat, aes(x = time, y = y, color = treatment)) +
    geom_point(show.legend = FALSE) +
    labs(y = 'response')+
    geom_line(aes(y = predict(sim_dat$fit_anova), 
                  group = subject, 
                  size = "Subjects"),
              show.legend = FALSE) +
    geom_line(data = sim_dat$pred_dat,
              aes(y = predict(sim_dat$fit_anova,
                              level = 0,
                              newdata = sim_dat$pred_dat),
                  size = "Population"),
              show.legend=FALSE) +
    guides(color = guide_legend(override.aes = list(size = 2))) +
    scale_size_manual(name = "Predictions",
                      values=c("Subjects" = 0.5, "Population" = 3)) +
    theme_classic() +
    theme(plot.title = element_text(size = txt, face = "bold"),
          text = element_text(size = txt)) +
    thm1

  #plot the LMEM predictions
  p5 <- ggplot(sim_dat$dat, aes(x = time, y = y, color = treatment)) +
    geom_point()+
    labs(y = 'response')+
    geom_line(aes(y = predict(sim_dat$fit_lme),
                  group = subject, size = "Subjects")) +
    geom_line(data = sim_dat$pred_dat,
              aes(y = predict(sim_dat$fit_lme, 
                              level = 0,
                              newdata = sim_dat$pred_dat),
                  size = "Population")) +
    guides(color = guide_legend(override.aes = list(size = 2)))+
    scale_size_manual(name = "Predictions",
                      values=c("Subjects" = 0.5, "Population" = 3)) +
    theme_classic() +
    theme(plot.title = element_text(size = txt, face = "bold"),
          text = element_text(size=txt))+
    thm1

  if(option=='simple'){
    return((p1 + p4 + p5) + plot_layout(nrow = 1) + 
             plot_annotation(tag_levels = 'A'))
  }
  else {
    return((p1 + p3 + p2 + p4 + p5) + plot_layout(nrow = 1) + 
             plot_annotation(tag_levels = 'A'))
  }
}
\end{lstlisting}

@@

The chunk below sources both \passthrough{\lstinline!example.R!} and \passthrough{\lstinline!plot\_example\_Appendix.R!} to simulate data and create the composite plots.

\begin{lstlisting}[language=R]
source(here::here("Manuscripts", "Manuscript_by_chapters-SIM_Revisions_final", 
                  "scripts", "example.R"))
source(here::here("Manuscripts", "Manuscript_by_chapters-SIM_Revisions_final", 
                  "scripts", "plot_example.R"))

A1<-plot_example(example(fun_type = "linear", error_type = "correlated"),
                 option = 'composite')

B1<-plot_example(example(fun_type = "linear", error_type = "independent"),
                 option = 'composite')

C1<-plot_example(example(fun_type = "quadratic", error_type = "correlated"), 
                 option = 'composite')

D1<-plot_example(example(fun_type = "quadratic", error_type = "independent"), 
                 option = 'composite') 
\end{lstlisting}

@@

\begin{lstlisting}[language=R]
#basis functions: this script creates Fig 2. by calculating a GAM for the Group 1 data of Figure 1,
# extracting the basis functions
#and creates objects p11,p12,p13,p14 for plotting, which are combined in b_plot to create the
#final composite figure
n_time = 6
x <- seq(1,6, length.out = n_time)
mu <- matrix(0, length(x), 2)
mu[, 1] <- -(0.25 * x^2) + 1.5 * x - 1.25 #mean response
mu[, 2] <- (0.25 * x^2) - 1.5 * x + 1.25 #mean response
y <- array(0, dim = c(length(x), 2, 10))
errors <- array(0, dim = c(length(x), 2, 10))
for (i in 1:2) {     # number of treatments
  for (j in 1:10) {  # number of subjects
    # compound symmetry errors
    errors[, i, j] <- rmvn(1, rep(0, length(x)), 0.1 * diag(6) + 0.25 * matrix(1, 6, 6))
    y[, i, j] <- mu[, i] + errors[, i, j]
  }
}

@@

smooth <- basis_plot %>%
  group_by(time) %>%
  summarize(smooth = sum(mod_val))

#original basis
sz <- 1
p11 <- ggplot(basis_plot,
              aes(x = time, y = value_orig, colour = as.factor(Basis))) +
  geom_line(size=sz, show.legend = FALSE) +
  geom_point(size = sz + 1, show.legend = FALSE) +
  labs(y = 'Basis functions') +
  facet_wrap(~ Basis, labeller = as_labeller(basis_names)) +
  theme_classic() +
  thm1

#penalized basis
p12 <- ggplot(basis_plot,
              aes(x = time, y = mod_val, colour = as.factor(Basis))) +
  geom_line(show.legend = FALSE, size = sz) +
  geom_point(show.legend = FALSE, size = sz + 1) +
  labs(y = 'Penalized \n basis functions') +
  scale_y_continuous(breaks = seq(-1, 1, 1)) +
  facet_wrap(~ Basis, labeller = as_labeller(basis_names)) +
  theme_classic() +
  thm1

#heatmap of the  coefficients
x_labels <- c("Intercept", "1", "2", "3", "4")
p13 <- ggplot(basis_plot,
              aes(x = Basis, y = Basis)) +
  geom_tile(aes(fill = coef), colour = "black") +
  scale_fill_gradient(low = "white", high = "#B50A2AFF") +
  labs(x = 'Basis', y = 'Basis') +
  scale_x_discrete(labels = x_labels) +
  geom_text(aes(label = round(coef, 2)),
            size = 7, show.legend = FALSE) +
  theme_classic()+
  theme(legend.title = element_blank())

#plotting simulated datapoints and smooth term
p14<-ggplot(data = dat,
            aes(x = time, y = y)) +
  geom_point(size = sz + 1, alpha = 0.5) +
  thm1 +
  labs(y = 'Simulated \n response') +
  geom_line(data = smooth, aes(x = time, y = smooth),
            color = "#6C581DFF", size = sz + 1)+
  theme_classic()

#Combining all
b_plot <- p11 + p13 + p12 + p14 + plot_annotation(tag_levels = 'A') &
  theme(text = element_text(size=18))
\end{lstlisting}

@@

The next chunk presents the code of the function \passthrough{\lstinline!simulate\_data()!} from the file \passthrough{\lstinline!simulate\_data.R!} in the \passthrough{\lstinline!scripts/!} folder, which we use in the main manuscript to simulate data that follows the reported trends in \(\mbox{StO}_2\) in the literature. Although the default standard deviation (SD) to generate the normally-distributed data in the function is 10\%, we use a standard deviation of 10\% in the main manuscript. We use the default number of 10 observations at each time point in the simulated data. Note that the input for the function is \passthrough{\lstinline!dat!}, needs to be a dataframe that contains the original values for the trends in each group over time from the literature. In our manuscript, we use the dataframe

\begin{lstlisting}[language=R]
#data from Vishwanath et. al.
dat <- tibble(StO2 = c(4, 27, 3, 2, 0.5, 7, 4, 50, 45, 56),
              Day = rep(c(0, 2, 5, 7, 10), times = 2),
              Group = as.factor(rep(c("Control", "Treatment"), each = 5)))
\end{lstlisting}

@@

\begin{lstlisting}[language=R]
#the same model used for the full dataset
mod_m1 <- gam(StO2_sim ~ Group+s(Day, by = Group, k = 5), 
              data  = dat_missing)
#appraise the model
appraise(mod_m1)

@@

f6 <- ggplot(ci, aes(x = Day, y = est, group = smooth)) +
  geom_line(lwd = 1) +
  geom_ribbon(data = ci, 
              aes(ymin = lower, ymax = upper, x = Day, 
                  group = smooth, fill = Group),
              inherit.aes = FALSE, alpha = 0.7, show.legend = FALSE) +
  geom_ribbon(data = si,
              aes(ymin = lower, ymax = upper, x = Day, 
                  group = smooth, fill =Group),
              inherit.aes = FALSE, alpha = 0.4, show.legend = TRUE) +
  geom_line(data = si, aes(Day, upper, color = Group), 
            size = 0.8, alpha = 0.7) +
  geom_point(data = dat_missing, aes(x = Day, y = StO2_sim, color=Group), 
             size = 1.5, alpha = 0.6, inherit.aes = FALSE) +
  labs(y = expression(atop(StO[2], '(incomplete observations)'))) +
  scale_x_continuous(breaks = c(0, 2, 5, 7, 10)) +
  theme_classic() +
  theme(axis.text=element_text(size = 22)) +
  thm1
\end{lstlisting}

@@

Then, \passthrough{\lstinline!difference\_smooths()!} function from the file \passthrough{\lstinline!difference\_smooths.R!} in the \passthrough{\lstinline!scripts/!} folder takes the output from the previous function and uses it to compute the simultaneous CI.
We present the code below.

@@

\begin{lstlisting}[language=R]
#Function that uses the source code of `gam.check` to obtain the estimates without the plots. The source can be checked by typing `gam.check` in the console.
gam_diagnostics <- function (b, old.style = FALSE, 
                           type = c("deviance", "pearson", "response"), 
                           k.sample = 5000, k.rep = 200, rep = 0,
                           level = 0.9, rl.col = 2, rep.col = "gray80", 
                           ...) {
  type <- match.arg(type)
  resid <- residuals(b, type = type)
  linpred <- if (is.matrix(b$linear.predictors) && !is.matrix(resid))
    napredict(b$na.action, b$linear.predictors[, 1])
  else napredict(b$na.action, b$linear.predictors)

  fv <- if (inherits(b$family, "extended.family"))
    predict(b, type = "response")
  else fitted(b)
  if (is.matrix(fv) && !is.matrix(b$y))
    fv <- fv[, 1]
  gamm <- !(b$method %in% c("GCV", "GACV", "UBRE",
                            "REML", "ML", "P-ML", "P-REML",
                            "fREML"))
  if (gamm) {
    message("\n'gamm' based fit - care required with interpretation.")
    message("\nChecks based on working residuals may be misleading.")
  }
  else {
    message("\nMethod:", b$method, "  Optimizer:", b$optimizer)
    if (!is.null(b$outer.info)) {
      if (b$optimizer[2] %in% c("newton", "bfgs")) {
        boi <- b$outer.info
        message("\n", boi$conv, " after ", boi$iter, " iteration", sep = "")
        if (boi$iter == 1)
          message(".")
        else message("s.")
        message("\nGradient range [", min(boi$grad), ",", max(boi$grad), "]", sep = "")
        message("\n(score ", b$gcv.ubre, " & scale ", b$sig2, ").", sep = "")
        ev <- eigen(boi$hess)$values
        if (min(ev) > 0)
          message("\nHessian positive definite, ")
        else message("\n")
        message("eigenvalue range [", min(ev), ",", max(ev), "].\n", sep = "")
      }
      else {
        message("\n")
        print(b$outer.info)
      }
    }
    else {
      if (length(b$sp) == 0)
        message("\nModel required no smoothing parameter selection")
      else {
        message("\nSmoothing parameter selection converged after",
            b$mgcv.conv$iter, "iteration")
        if (b$mgcv.conv$iter > 1)
          message("s")
        if (!b$mgcv.conv$fully.converged)
          message(" by steepest\ndescent step failure.\n")
        else message(".\n")
        message("The RMS", b$method, "score gradient at convergence was",
            b$mgcv.conv$rms.grad, ".\n")
        if (b$mgcv.conv$hess.pos.def)
          message("The Hessian was positive definite.\n")
        else message("The Hessian was not positive definite.\n")
      }
    }
    if (!is.null(b$rank)) {
      message("Model rank = ", b$rank, "/", length(b$coefficients), "\n")
    }
  }
  message("\n")
  kchck <- k.check(b, subsample = k.sample, n.rep = k.rep)
  if (!is.null(kchck)) {
    message("Basis dimension (k) checking results. Low p-value (k-index<1) may\n")
    message("indicate that k is too low, especially if edf is close to k'.\n\n")
    printCoefmat(kchck, digits = 3)
  }
}
\end{lstlisting}
